Psychophysics of musical elements in the discrete-time representation of sound

### 2013-10-02 article/article.tex
ateh os ruidos

For filters with other frequency responses, it is possible to apply them successively. Another possibility is to use a biquad 'filter receipt'\footnote{Short for 'biquadratic': its transfer function has two poles and two zeros, i.e. its first direct form consists of two quadratic polynomials forming the fraction: $\mathbb{H}(z)=\frac{a_0+a_1.z^{-1}+a_2.x^{-2}}{1- b_1.z^{-1} -b_2 . z^{-2}}$.} or the calculation of Chebichev filter coefficients\footnote{Butterworth and Elliptical filters can be considered as special cases of Chebichev filters~\cite{Openheim,smith}.}.
Both alternatives are explored by referenced studies, in special~\cite{JOSFM,smith}, and by the collection of filters maintained by the \emph{Music-DSP} community of the Columbia University .\cite{music-dsp,Openheim}

\begin{enumerate}
  \item Low-pass with simple pole and module of the frequency response in the upper left corner of Figure~\ref{fig:iir}. The general equation has as reference by the cutoff frequency $f_c \in (0,\frac{1}{2})$, fraction of the sample frequency $f_a$ that there is approximately an attenuation of $3dB$.
  The coefficients $a_0$ and $b_1$ of the IIR filter are given by means of the intermediate variable $x \in [e^{-\pi},1]$:

@@

@@

  \item High-pass filter with simple pole and modules of its frequency responses at the upper right corner of Figure~\ref{fig:iir}. The general equation with cutoff frequency $f_c \in (0,\frac{1}{2})$ is calculated by means of the intermediate variable $x \in [e^{-\pi},1]$:

@@

\item Notch filter. This filter is parametrized by a center frequency\footnote{Attention with the cutoff frequency $f_c$ in low-pass and high-pass filters.} $f_c$ and band width $bw$ -- $f_c \pm bw$ that results in $0.707$ of the amplitude, i.e. attenuation of $3dB$ -- both given as fractions of $f_a$, therefore $f,\; bw \in (0,0.5)$.

In order to simplify, consider the auxiliary variables $K$ and $R$ defined as:

@@

The band-pass filter in the lower left corner of Figure~\ref{fig:iir} has the following coefficients in equation~\ref{eq:diferencas}:

@@

The coefficients of band-reject filter are:

@@

\noindent with the module of its frequency response represented in the lower left corner of the Figure~\ref{fig:iir}.

@@

Generally, sounds without a defined pitch are called noise~\cite{Lacerda}.
They are important parts of the musical sounds with undefined pitch like the noise present in piano notes, violin, etc. Besides that, the majority of percussion instruments do not have defined pitch and their sounds are generally comprehended as noise~\cite{Roederer}. In electronic music, including electro-acoustic and dance genres, noise has diversified use, and commonly characterizes the music style~\cite{Cook}.

>>>>

For filters with other frequency responses, it is possible to apply them successively. Another possibility is to use a biquad 'filter receipt'\footnote{Short for 'biquadratic': its transfer function has two poles and two zeros, i.e. its first direct form consists of two quadratic polynomials forming the fraction: $\mathbb{H}(z)=\frac{a_0+a_1.z^{-1}+a_2.x^{-2}}{1- b_1.z^{-1} -b_2 . z^{-2}}$.} or the calculation of Chebichev filter coefficients\footnote{Butterworth and Elliptical filters can be considered as special cases of Chebichev filters~\cite{Openheim,smith}.}.
Both alternatives are explored by referenced studies, in special~\cite{JOSFM,smith}, and by the collection of filters maintained by the \emph{Music-DSP} community of the Columbia University~\cite{music-dsp,Openheim}.

@@

\begin{enumerate}
  \item Low-pass with a simple pole, module of the frequency response in the upper left corner of figure~\ref{fig:iir}. The general equation has the cutoff frequency $f_c \in (0,\frac{1}{2})$, fraction of the sample frequency $f_s$ in which occurs an attenuation of $3dB$.
  The coefficients $a_0$ and $b_1$ of the IIR filter are given by means of the intermediate variable $x \in [e^{-\pi},1]$:

@@

  \item High-pass filter with a simple pole,  module of its frequency responses at the upper right corner of figure~\ref{fig:iir}. The general equation with cutoff frequency $f_c \in (0,\frac{1}{2})$ is calculated by means of the intermediate variable $x \in [e^{-\pi},1]$:

@@

\item Notch filter. This filter is parametrized by a center frequency $f_c$ and bandwidth $bw$, both given as fraction of $f_s$, therefore $f,\; bw \in (0,0.5)$. Both frequencies $f_c \pm bw$ has $\approx 0.707$ of the amplitude, i.e. attenuation of $3dB$.
Auxiliary variables $K$ and $R$ are defined as:

@@

The band-pass filter in the lower left corner of figure~\ref{fig:iir} has the following coefficients:

@@

The coefficients of band-reject filter, depicted in the lower right of figure~\ref{fig:iir}, are:

@@

@@

Generally, sounds without a defined pitch are called noise~\cite{Lacerda}.
They are important musical sounds, as noise present in piano notes, violin, etc. Besides that, the majority of percussion instruments do not have precise pitch and their sounds are generally comprehended as noise~\cite{Roederer}. In electronic music, including electro-acoustic and dance genres, noise has diverse uses, and frequently characterizes the music style~\cite{Cook}.
